---
title: "Model Selection"
author: "Nathan Kolbow"
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = FALSE)
library(tidyverse)
library(ggplot2)
library(patchwork)
library(ggh4x)
library(cowplot)
```

```{r load_data, include=FALSE}
alphas <- c(-0.001, 0, 0.001, 0.01, 0.05, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.99, 0.99999, 1.0, 1.01)
st <- Sys.time()
alldata <- rbind(
  read.csv("dat-eps0.0.csv"),
  read.csv("dat-eps0.1.csv"),
  read.csv("dat-eps0.01.csv"),
  read.csv("dat-eps0.001.csv"),
  # read.csv("dat-eps0.0001.csv"),
  read.csv("dat-eps1.0.csv")
) %>%
  #filter(test != "DDSE" & test != "Djump") %>%
  mutate(eps = if_else(model == "quartet", 0.0, eps))

df <- alldata %>%
  filter(test != "CLIC" & test != "DDSE" & test != "Djump") %>%
  crossing(alphas) %>%
  rename(alpha = alphas) %>%
  arrange(simid, esth, alpha) %>%
  group_by(test, simid, model, eps, alpha) %>%
  mutate(
    cutoff = alpha, #1 - (1-alpha)^(esth),
    rejectH0 = if_else(esth == 0, TRUE, result < cutoff),
    first_failure_to_reject = (!rejectH0 & row_number() == min(c(row_number()[!rejectH0], max(row_number())))),
    selected = lead(first_failure_to_reject, default = FALSE),
    selected = if_else(rep(all(rejectH0), n()), row_number() == max(row_number()), selected),
  )

df <- df %>%
  rbind(
    alldata %>%
      filter(test == "CLIC" | test == "DDSE" | test == "Djump") %>%
      group_by(test, simid, model, eps) %>%
      mutate(
        alpha = 0, cutoff = 0, rejectH0 = FALSE, first_failure_to_reject = FALSE, selected = result == min(result)
      )
  )

best_alpha <- df %>%
  filter(selected) %>%
  group_by(test, model, eps, alpha, trueh, ngt) %>%
  summarise(acc = mean(trueh == esth), .groups = "drop_last") %>%
  group_by(test, model, eps, alpha, ngt) %>%
  summarise(acc = prod(0.1 + acc) - 0.1, .groups = "drop_last") %>%
  group_by(test, model, eps, ngt) %>%
  slice_max(acc, n=1, with_ties=FALSE)

df <- df %>%
  semi_join(best_alpha, c("test", "model", "eps", "alpha", "ngt")) %>%
  mutate(result = if_else(test == "CLIC", sign(result) * log(0.1 + abs(result)), result))
Sys.time() - st    # 37.83776
```



::: {.alert .alert-info}
**Note:** Code chunk `include` default value has been set to `FALSE`, so code chunks (and their plots) will not show unless explicitly set with `include=TRUE`.
:::

<!-- Comparing all tests -->

```{r}
modelacc_df <- df %>%
  filter(selected) %>%
  group_by(test, model, eps) %>%
  summarise(
    acc = mean(trueh == esth),
    accH0 = mean(esth[trueh == 0] == 0),
    accHnonzero = mean(esth[trueh != 0] != 0)
  )
modelacc_df
df %>%
  filter(eps == 1 & trueh == 2 & model == "joint" & test == "cwP")
```

<!-- All eps=EPS -->

```{r}
EPS <- 0.
df %>%
  filter(eps == EPS & test == "CLIC") %>%
  ggplot(aes(x = esth, y = result, color = factor(trueh), group = interaction(simid, test, model, ngt))) +
  geom_point() +
  geom_line() +
  facet_nested(ngt ~ model, scales="free") +
  geom_hline(yintercept = 0, alpha = 0.25, linetype = "dashed")
```

# Examining the tests

## Tests by $\epsilon$

### $\epsilon=0$

```{r, fig.width=10, fig.height=5, include=TRUE}
EPS <- 0
df %>%
  filter(eps == EPS & selected) %>%
  group_by(test, model, trueh, ngt) %>%
  summarise(acc = if_else(n() == 0, -1, mean(trueh == esth)), .groups = "drop") %>%
  ggplot(aes(x = factor(trueh), y = acc, fill = test)) +
  geom_col(position = position_dodge2(), color = "black") +
  facet_nested(test ~ ngt + model)
```

### $\epsilon=0.1$

```{r, fig.width=10, fig.height=5, include=TRUE}
EPS <- 0.1
df %>%
  filter(eps == EPS & selected) %>%
  group_by(test, model, trueh, ngt) %>%
  summarise(acc = if_else(n() == 0, -1, mean(trueh == esth)), .groups = "drop") %>%
  ggplot(aes(x = factor(trueh), y = acc, fill = test)) +
  geom_col(position = position_dodge2(), color = "black") +
  facet_nested(test ~ ngt + model)
```

### $\epsilon=0.01$

```{r, fig.width=10, fig.height=5, include=TRUE}
EPS <- 0.01
df %>%
  filter(eps == EPS & selected) %>%
  group_by(test, model, trueh, ngt) %>%
  summarise(acc = if_else(n() == 0, -1, mean(trueh == esth)), .groups = "drop") %>%
  ggplot(aes(x = factor(trueh), y = acc, fill = test)) +
  geom_col(position = position_dodge2(), color = "black") +
  facet_nested(test ~ ngt + model)
```

### $\epsilon=1$

```{r, fig.width=10, fig.height=5, include=TRUE}
EPS <- 1
df %>%
  filter(eps == EPS & selected) %>%
  group_by(test, model, trueh, ngt) %>%
  summarise(acc = if_else(n() == 0, -1, mean(trueh == esth)), .groups = "drop") %>%
  ggplot(aes(x = factor(trueh), y = acc, fill = test)) +
  geom_col(position = position_dodge2(), color = "black") +
  facet_nested(test ~ ngt + model)
```






## Comparing $\epsilon$ values for each test

```{r, fig.width=12, fig.height=6, include=TRUE}
compare_eps_plot <- function(tib, testname) {
  tib %>%
    filter(selected & test == testname) %>%
    group_by(eps, ngt, trueh, model) %>%
    summarise(acc = mean(trueh == esth), .groups = "drop") %>%
    mutate(trueh = factor(trueh), eps = factor(eps)) %>%
    ggplot(aes(x = eps, y = acc, color = trueh, group = interaction(model, ngt, trueh))) +
    geom_point() +
    geom_line() +
    facet_nested(ngt ~ model, scales = "free") +
    theme_bw() +
    ggtitle(paste0("Test: ", testname)) +
    labs(
      x = bquote(epsilon),
      y = "Accuracy",
      color = "True H"
    )
}

compare_eps_plot(df, "cLR") + compare_eps_plot(df, "cLRI") + plot_layout(axes = "collect", guides = "collect")
compare_eps_plot(df, "cLR1") + compare_eps_plot(df, "cLR2") + plot_layout(axes = "collect", guides = "collect")
(compare_eps_plot(df, "cw") + compare_eps_plot(df, "cwP") + compare_eps_plot(df, "CLIC")) + plot_layout(axes = "collect", guides = "collect")
```

## Over marginalized comparisons

```{r, fig.width=12, fig.height=5, include=TRUE}
sel_df <- filter(df, selected) %>% ungroup() %>% mutate(trueh = factor(trueh), ngt = factor(ngt), eps = factor(eps))

bests_df <- sel_df %>%
  group_by(test, trueh, ngt, model, eps) %>%
  summarise(acc = mean(esth == trueh), .groups = "drop")

# Test
semi_join(
  sel_df,
  bests_df %>%
    group_by(test) %>%
    summarise(acc = mean(acc), .groups = "drop"),
  "test"
) %>% 
  group_by(ngt, test) %>%
  summarise(acc = mean(trueh == esth)) %>%
  ggplot(aes(x = ngt, y = acc, fill = test)) +
  geom_col(position = position_dodge2(), color = "black") +
  facet_grid(~ test) +
  theme_bw() +
  labs(
    x = "Number of loci",
    y = "Accuracy"
  ) +
  guides(fill = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.ticks.length = unit(0, "lines"))

# Model
semi_join(
  sel_df,
  bests_df %>%
    group_by(model) %>%
    summarise(acc = mean(acc), .groups = "drop"),
  "model"
) %>% 
  group_by(ngt, model) %>%
  summarise(acc = mean(trueh == esth)) %>%
  ggplot(aes(x = ngt, y = acc, fill = model)) +
  geom_col(position = position_dodge2(), color = "black") +
  facet_grid(~ model) +
  theme_bw() +
  labs(
    x = "Number of loci",
    y = "Accuracy"
  ) +
  guides(fill = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.ticks.length = unit(0, "lines"))

# Epsilon
semi_join(
  sel_df,
  bests_df %>%
    group_by(eps) %>%
    summarise(acc = mean(acc), .groups = "drop"),
  "eps"
) %>% 
  group_by(ngt, eps) %>%
  summarise(acc = mean(trueh == esth)) %>%
  ggplot(aes(x = ngt, y = acc, fill = eps)) +
  geom_col(position = position_dodge2(), color = "black") +
  facet_grid(~ eps) +
  theme_bw() +
  labs(
    x = "Number of loci",
    y = "Accuracy"
  ) +
  guides(fill = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.ticks.length = unit(0, "lines"))
```

# Model + $\epsilon$

```{r, fig.width=8, fig.height=3, include=TRUE}
subdf <- semi_join(
  sel_df,
  bests_df %>%
    group_by(model, eps) %>%
    summarise(acc = mean(acc), .groups = "drop"),
  c("model", "eps")
) %>%
  mutate(eps = paste0("eps=", eps)) %>%
  filter(model != "quartet" | eps == "eps=0")

pjoint <- subdf %>%
  filter(model == "joint") %>%
  group_by(ngt, model, eps) %>%
  summarise(acc = mean(trueh == esth), .groups = "drop") %>%
  ggplot(aes(x = ngt, y = acc, fill = model)) +
  geom_col(position = position_dodge2(), color = "black", fill = "#e41a1c") +
  facet_grid( ~ eps) +
  theme_bw() +
  labs(
    x = "Number of loci",
    y = "Accuracy"
  ) +
  guides(fill = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.ticks.length = unit(0, "lines")) +
  ggtitle("Joint") +
  lims(y = c(0, 1))
pmarg <- subdf %>%
  filter(model == "marginal") %>%
  group_by(ngt, model, eps) %>%
  summarise(acc = mean(trueh == esth), .groups = "drop") %>%
  ggplot(aes(x = ngt, y = acc, fill = model)) +
  geom_col(position = position_dodge2(), color = "black", fill = "#377eb8") +
  facet_grid( ~ eps) +
  theme_bw() +
  labs(
    x = "Number of loci",
    y = "Accuracy"
  ) +
  guides(fill = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.ticks.length = unit(0, "lines")) +
  ggtitle("Marginal") +
  lims(y = c(0, 1))
pquart <- subdf %>%
  filter(model == "quartet") %>%
  group_by(ngt, model, eps) %>%
  summarise(acc = mean(trueh == esth), .groups = "drop") %>%
  ggplot(aes(x = ngt, y = acc, fill = model)) +
  geom_col(position = position_dodge2(), color = "black", fill = "#4daf4a") +
  theme_bw() +
  facet_grid( ~ eps) +
  labs(
    x = "Number of loci",
    y = "Accuracy"
  ) +
  guides(fill = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.ticks.length = unit(0, "lines")) +
  ggtitle("Quartet") +
  lims(y = c(0, 1))

pjoint
pmarg
pquart
```


# Determining the best tests

```{r, fig.width=6, fig.height=4, include=TRUE}
best_test_df <- df %>%
  filter(selected) %>%
  group_by(test, model, eps, trueh) %>%
  summarise(
    acc = mean(trueh == esth),
    meandiff = mean(abs(trueh - esth)),
    ntests = n()
  ) %>%
  group_by(test, model, eps) %>%
  summarise(
    acc = prod(acc),
    ntests = sum(ntests)
  ) %>%
  ungroup() %>%
  filter(acc == max(acc))

semi_join(df, best_test_df, c("test", "model", "eps")) %>%
  filter(selected) %>%
  group_by(ngt, trueh, test) %>%
  summarise(acc = mean(trueh == esth)) %>%
  ggplot(aes(x = factor(trueh), y = acc, fill = factor(trueh), group = interaction(test, trueh, ngt))) +
  geom_col(position = position_dodge2(), color = "black") +
  theme_bw() +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  facet_wrap(~ ngt) +
  ggtitle(paste0(
    best_test_df$test, " [", best_test_df$model, "] (eps=", best_test_df$eps, ")", collapse="\n"
  ))
```

Further segregating by `ngt`.

```{r, fig.width=10, fig.height=6, include=TRUE}
best_test_df <- df %>%
  filter(selected) %>%
  group_by(test, model, eps, trueh, ngt) %>%
  summarise(
    acc = mean(trueh == esth),
    meandiff = mean(abs(trueh - esth)),
    ntests = n()
  ) %>%
  group_by(test, model, eps, ngt) %>%
  summarise(
    acc = prod(acc),
    ntests = sum(ntests)
  ) %>%
  group_by(ngt) %>%
  filter(acc == max(acc)) %>%
  ungroup()

p_ngtdf <- df %>%
  semi_join(best_test_df, c("test", "model", "eps", "ngt")) %>%
  filter(selected) %>%
  group_by(ngt, trueh, test) %>%
  summarise(acc = mean(trueh == esth))

p10 <- ggplot(filter(p_ngtdf, ngt == 10), aes(x = factor(trueh), y = acc, fill = factor(trueh), group = interaction(test, trueh, ngt))) +
  geom_col(position = position_dodge2(), color = "black") +
  theme_bw() +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  facet_wrap(~ ngt) +
  ggtitle(paste0(
    first(filter(best_test_df, ngt == 10)$test), " [", first(filter(best_test_df, ngt == 10)$model), "] (eps=", first(filter(best_test_df, ngt == 10)$eps), ")", "[", nrow(filter(best_test_df, ngt == 10)), "]", collapse="\n"
  ))
p50 <- ggplot(filter(p_ngtdf, ngt == 50), aes(x = factor(trueh), y = acc, fill = factor(trueh), group = interaction(test, trueh, ngt))) +
  geom_col(position = position_dodge2(), color = "black") +
  theme_bw() +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  facet_wrap(~ ngt) +
  ggtitle(paste0(
    first(filter(best_test_df, ngt == 50)$test), " [", first(filter(best_test_df, ngt == 50)$model), "] (eps=", first(filter(best_test_df, ngt == 50)$eps), ")", "[", nrow(filter(best_test_df, ngt == 50)), "]", collapse="\n"
  ))
p100 <- ggplot(filter(p_ngtdf, ngt == 100), aes(x = factor(trueh), y = acc, fill = factor(trueh), group = interaction(test, trueh, ngt))) +
  geom_col(position = position_dodge2(), color = "black") +
  theme_bw() +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  facet_wrap(~ ngt) +
  ggtitle(paste0(
    first(filter(best_test_df, ngt == 100)$test), " [", first(filter(best_test_df, ngt == 100)$model), "] (eps=", first(filter(best_test_df, ngt == 100)$eps), ")", "[", nrow(filter(best_test_df, ngt == 100)), "]", collapse="\n"
  ))
p500 <- ggplot(filter(p_ngtdf, ngt == 500), aes(x = factor(trueh), y = acc, fill = factor(trueh), group = interaction(test, trueh, ngt))) +
  geom_col(position = position_dodge2(), color = "black") +
  theme_bw() +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  facet_wrap(~ ngt) +
  ggtitle(paste0(
    first(filter(best_test_df, ngt == 500)$test), " [", first(filter(best_test_df, ngt == 500)$model), "] (eps=", first(filter(best_test_df, ngt == 500)$eps), ")", "[", nrow(filter(best_test_df, ngt == 500)), "]", collapse="\n"
  ))
p1000 <- ggplot(filter(p_ngtdf, ngt == 1000), aes(x = factor(trueh), y = acc, fill = factor(trueh), group = interaction(test, trueh, ngt))) +
  geom_col(position = position_dodge2(), color = "black") +
  theme_bw() +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  facet_wrap(~ ngt) +
  ggtitle(paste0(
    first(filter(best_test_df, ngt == 1000)$test), " [", first(filter(best_test_df, ngt == 1000)$model), "] (eps=", first(filter(best_test_df, ngt == 1000)$eps), ")", "[", nrow(filter(best_test_df, ngt == 1000)), "]", collapse="\n"
  ))

p10 + p50 + p100 + p500 + p1000 + plot_layout(axes="collect", guides="collect")
```

Results based on the tests that are tied for best performance with 50 gene trees (because >50 has many contenders getting perfect scores)

```{r, fig.height=6, fig.width=10, include=TRUE}
best_test_df <- df %>%
  filter(selected & ngt == 50) %>%
  group_by(test, model, eps, trueh) %>%
  summarise(
    acc = mean(trueh == esth),
    meandiff = mean(abs(trueh - esth)),
    ntests = n(),
    .groups = "drop"
  ) %>%
  group_by(test, model, eps) %>%
  summarise(
    prodacc = prod(acc),
    ntests = sum(ntests),
    .groups = "drop"
  ) %>%
  filter(prodacc == max(prodacc)) %>%
  ungroup()

plots <- list()
for(i in 1:nrow(best_test_df)) {
  plots[[i]] <- semi_join(df, best_test_df[i, ], c("test", "model", "eps")) %>%
    filter(selected) %>%
    group_by(ngt, trueh, test) %>%
    summarise(acc = mean(trueh == esth), .groups = "drop") %>%
    ggplot(aes(x = factor(trueh), y = acc, fill = factor(trueh), group = interaction(test, trueh, ngt))) +
    geom_col(position = position_dodge2(), color = "black") +
    theme_bw() +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
    facet_wrap(~ ngt) +
    ggtitle(paste0(
      best_test_df$test[i], " [", best_test_df$model[i], "] (eps=", best_test_df$eps[i], ")", collapse="\n"
    ))
}
plots
```


# DDSE / Djump




































